---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggpointgrid

This package provides some simple geoms (`geom_pointgrid`, `geom_pointrect`) derived from geom_point to rearrange scatterplot coordinates on regular grids while strictly avoiding over-plotting. The applications might be similar to `geom_jitter`.

### Installation

You can install the development version from github with the following command (in your R console):

```
if(!require('remotes')) install.packages('remotes')
remotes::install_github("nevrome/ggpointgrid")
```

### Example

```{r}
library(ggplot2)
library(magrittr)
library(palmerpenguins)
```

```{r, warning=FALSE}
palmerpenguins::penguins %>%
  ggplot() +
  ggpointgrid::geom_pointgrid(
    aes(x = body_mass_g, y = bill_length_mm, shape = species, color = sex),
    grid_x = 50,
    grid_y = 50
  )

palmerpenguins::penguins %>%
  dplyr::arrange(island) %>%
  ggplot() +
  ggpointgrid::geom_pointrect(
    aes(x = sex, y = species, color = island), 
    scale_x = 0.05, 
    scale_y = 0.05
  )
```

