---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# ggpointgrid

This package provides some simple geoms derived to rearrange scatterplot coordinates on regular grids while strictly avoiding over-plotting. The applications are similar to `geom_jitter`.

### Installation

You can install the development version from github with the following command (in your R console):

```
if(!require('remotes')) install.packages('remotes')
remotes::install_github("nevrome/ggpointgrid")
```

### Example

```{r}
library(ggplot2)
set.seed(5)
```

#### `geom_pointgrid`

```{r, warning=FALSE}
test_data_1 <- tibble::tibble(
  x = rep(c(1,1,2,3,3), times = 10),
  y = rep(c(1,3,2,1,3), times = 10),
  var = sample(c("A", "B", "C"), size = 50, replace = T)
)

p1 <- test_data_1 |>
  ggplot() +
  geom_point(aes(x, y, color = var)) +
  coord_fixed()

p2 <- test_data_1 |>
  ggplot() +
  geom_jitter(aes(x, y, color = var), width = 0.3, height = 0.3) +
  coord_fixed()

p3 <- test_data_1 |>
  ggplot() +
  ggpointgrid::geom_pointgrid(
    aes(x, y, color = var),
    grid_x = 15,
    grid_y = 15
  ) +
  coord_fixed()

cowplot::plot_grid(p1, p2, p3)
```

`geom_textgrid` performs the same arrangement operation on text data, so can be used in place of `geom_text`.

#### `geom_pointrect`

```{r, warning=FALSE}
test_data_2 <- tibble::tibble(
  x = rep(letters[c(1,1,2,3,3)], times = 10),
  y = rep(letters[c(1,3,2,1,3)], times = 10),
  var = sample(c("A", "B", "C"), size = 50, replace = T)
) |> dplyr::arrange(var)

p4 <- test_data_2 |>
  ggplot() +
  geom_point(aes(x, y, color = var)) +
  coord_fixed()

p5 <- test_data_2 |>
  ggplot() +
  geom_jitter(aes(x, y, color = var), width = 0.3, height = 0.3) +
  coord_fixed()

p6 <- test_data_2 |>
  ggplot() +
  ggpointgrid::geom_pointrect(
    aes(x, y, color = var)
  ) +
  coord_fixed()

cowplot::plot_grid(p4, p5, p6)
```
